---
title: "Fitness_CART"
author: "Ananya Mathur"
date: "10/20/2020"
output:
  pdf_document: default
  html_document: default
---

##Introduction
In this mini project, the aim is to analyze the data in the training set and classify it into A,B,C,D or E. The data contains a lot of fitness variables based on fitbit type devices and the goal is to classify them into catergories of exercise.

###Installing the packages and Libraries
caret and ggplot2 are the various packages that are required to install for importing the required libraries.

```{r}
#install.packages("caret")
#install.packages("ggplot2")
#install.packages("rpart")
#install.packages("randomForest")
```
#Importing Libraries
caret library i.e the classification and regression Trees are imported.
readr library is used for reading csv files in R.
```{r}
library(caret)
library(readr)
library(rpart)
library(randomForest)
```

## Getting the Data

```{r}
#Getting the Data
###reading the training dataset
pml_training <- read_csv("Data Science, John Hopkins/pml-training.csv")
summary(pml_training)
#reading the test dataset
pml_testing <- read_csv("Data Science, John Hopkins/pml-testing.csv")
summary(pml_testing)
```

##Exploratory Data Analysis

```{r}
#Exploratory Data Analysis
###for obtaining the type of each column
str(pml_training)

###for checking the number of rows and columns in the dataset, the train dataset contains 19622 rows and 160 columns.
dim(pml_training)

###taking a count for the final results
table(pml_training$classe)

```
#Cleaning the Data
Dealing with missing values


```{r}
#Exploratory Data Analysis
###for obtaining the type of each column

Count_na = sapply(1:dim(pml_training)[2],function(x)sum(is.na(pml_training[,x])))
List_na = which(Count_na>0)
colnames(pml_training[,c(1:7)])

```
###removing the null values from the training set

```{r}
pml_training = pml_training[,-List_na]
pml_training = pml_training[,-c(1:7)]
pml_training$classe = factor(pml_training$classe)
###removing the null values from the test set
pml_testing = pml_testing[,-List_na]
pml_testing = pml_testing[,-c(1:7)]
```
#Building the Model
Decision Tree
```{r}
#Exploratory Data Analysis
###for obtaining the type of each column

Classifier = rpart(formula = classe~.,data=pml_training)
x_pred = predict(Classifier,pml_testing)
print(x_pred)

x_pred_actual = predict(Classifier,pml_testing,type="class")
print(x_pred_actual)

###confusion Matrix
am = table(pml_testing[-1],x_pred_actual)
print(am)
```
#Random Forest

```{r}
#Exploratory Data Analysis
###for obtaining the type of each column

rfclassifier = randomForest(formula = classe~.,data=pml_training, ntree = 10)
x_pred_rf = predict(rfclassifier,pml_testing)
print(x_pred_rf)
```



